---
title: "Introduction to csSampling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to csSampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
author:
  - name: Ryan Hornby
    affiliation: |
      | Vassar College \AND
    address: |
      | Poughkeepsie, NY, USA
    # use this syntax to add text on several lines
    email: \email{rhornby@vassar.edu}
    url: https://github.com/RyanHornby
  - name: Matthew R. Williams
    address: |
      | VA, USA
    affiliation: |
      | RTI International
    email: \email{mrwilliams@rti.org}
    # To add another line, use \AND at the end of the previous one as above
  - name: Terrance D. Savitsky
    address: |
      | Washington, DC, USA
    affiliation: |
      | Office of Survey Methods Research
      | Bureau of Labor Statistics
    email: \email{Savitsky.Terrance@bls.gov}
  - name: Mahmoud Elkasabi
    address: |
      | MD, USA
    affiliation: |
      | RTI International
    email: \email{melkasabi@rti.org}
---

  We present csSampling, an R package for estimation of Bayesian models for data collected from complex survey samples. csSampling combines functionality from the probabilistic programming language Stan (via the rstan and brms R packages) and the survey R package. Under the proposed approach, the user creates a survey weighted model in brms or provides custom weighted model via rstan. Survey design information is provided via svydesign function of the survey package. The cs_sampling function of csSampling estimates the weighted stan model and provides an asymptotic covariance correction for model mis-specification due to using survey sampling weights as plug in values in the likelihood. This is often known as a design effect which is the ratio between the variance from complex survey sample and a simple random sample.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r echo=TRUE, message=FALSE, warning = FALSE}
library(csSampling)
library(rstan)
library(brms)
library(survey)
rstan_options(auto_write = TRUE)
```

## Example 1: continuous dependent variable

Survey Design Information

```{r, echo = TRUE, message = FALSE}
data(api)
apistrat$wt <- apistrat$pw /mean(apistrat$pw)

dstrat <-
  svydesign(id=~1,strata=~stype, weights=~wt, data=apistrat, fpc=~fpc)

```

Define and Run the Stan Model Via BRMS Wrapper

```{r echo = TRUE, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE}
set.seed(12345)
model_formula <- formula("api00|weights(wt) ~ 
                            ell + meals + mobility")
mod.brms <- cs_sampling_brms(svydes = dstrat, 
    brmsmod = brmsformula(model_formula, center = F), 
    data = apistrat, family = gaussian())
```

Plot the Results

```{r message=FALSE, warning=FALSE, cache=FALSE, include=FALSE}
plot(mod.brms)
```

```{r echo = TRUE, out.width = ".45\\linewidth", fig.show = "hold"}
plot(mod.brms)
plot(mod.brms, varnames = paste("b", 1:4, sep =""))
```

## Example 2: multinomial dependent variable

```{r echo = TRUE,message=FALSE, warning=FALSE, include=TRUE}
data(api)
apiclus1$wt <- apiclus1$pw/mean(apiclus1$pw)
dclus1<-svydesign(id=~dnum, weights=~wt, data=apiclus1, fpc=~fpc)
rclus1<-as.svrepdesign(dclus1)
svymean(~stype, rclus1)
```

Construct a Stan mode

```{r echo = TRUE, message=FALSE, warning=FALSE, include=TRUE, cache = TRUE}
mod_dm <- stan_model(model_code = load_wt_multi_model())
```

Prepare data for Stan modelling

```{r echo = TRUE, message=FALSE, warning=FALSE, include=TRUE}
#Set the Data for Stan
y <- as.factor(rclus1$variables$stype)
yM <- model.matrix(~y -1)
n <- dim(yM)[1]
K <- dim(yM)[2]
alpha<-rep(1,K)
weights <- rclus1$pweights
data_stan<-list("y"=yM,"alpha"=alpha,"K"=K, "n" = n, "weights" = weights)
ctrl_stan<-list("chains"=1,"iter"=2000,"warmup"=1000,"thin"=1)
```

Run the Stan mode

```{r echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE}
mod1 <- cs_sampling(svydes = rclus1, mod_stan = mod_dm, 
                    data_stan = data_stan, ctrl_stan = ctrl_stan, rep_design = T)
```

Plot the results

```{r echo = TRUE, out.width = ".45\\linewidth", fig.show = "hold"}
plot(mod1)
plot(mod1, varnames = paste("theta",1:3, sep = ""))
```

Compare Summary Statistics

```{r echo = TRUE}
#hybrid variance adjustedM2
cbind(colMeans(mod1$adjusted_parms[,4:6]), 
      sqrt(diag(cov(mod1$adjusted_parms[,4:6]))))
#Replication
svymean(~stype, rclus1)
```

## Example 3: Bernoulli-distributed dependent variable

```{r echo = TRUE, cache = TRUE}
dat14 <- csSampling::dat14
#subset to adults
dat14 <- dat14[as.numeric(dat14$CATAG6) > 1,]
#normalize weights to sum to sample size
dat14$WTS <- dat14$ANALWT_C/mean(dat14$ANALWT_C)
svy14 <- svydesign(ids = ~VEREP, strata = ~VESTR, 
                   weights = ~WTS, data = dat14, nest = T)
```

Construct a Stan model

```{r echo = TRUE}

model_formula <- formula("CIGMON|weights(WTS) ~ AMDEY2_U")
stancode <- make_stancode(brmsformula(model_formula,center = F), 
		data = dat14, family = bernoulli(), save_model = "brms_wt_log.stan")
mod_brms  <- stan_model('brms_wt_log.stan')
```

Prepare data for Stan modelling

```{r echo = TRUE}
data_brms <- make_standata(brmsformula(CIGMON|weights(WTS) ~ AMDEY2_U,center = FALSE), 
		data = dat14, family = bernoulli())
```

Run the Stan model

```{r, echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE}
set.seed(12345) #set seed to fix output for comparison
mod.brms <- cs_sampling(svydes = svy14, mod_stan = mod_brms, data_stan = data_brms)
```

Plot the results

```{r echo = TRUE, out.width = ".45\\linewidth", fig.show = "hold"}
plot(mod.brms, varnames = paste("b", 1:2, sep =""))

```
